### 2017-06-02 paper/figures/example-stats.tex

\begin{floattable}
\begin{deluxetable}{cccccccc}
\caption{The scaling and convergence stats for each example. \label{tab:example-stats}}
\tablehead{\colhead{example} & \colhead{$N$} & \colhead{$J$} & \colhead{direct cost} & \colhead{\celerite\ cost} & \colhead{$D$} & \colhead{model evaluations} & \colhead{effective samples}\\
&&& \colhead{ms} & \colhead{ms} &&&}
\startdata
1 & 200 & 1 & 2.85 & 0.26 & 3 & 80000 & 1737 \\
2 & 100 & 2 & 0.67 & 0.36 & 4 & 80000 & 1134 \\
3 & 6950 & 2 & 8119.11 & 1.47 & 4 & 176000 & 2900 \\
4 & 1440 & 8 & 828.93 & 2.74 & 9 & 640000 & 1443 \\
5 & 1000 & 2 & 82.73 & 0.94 & 13 & 640000 & 1141 \\
\enddata
\end{deluxetable}
\end{floattable}