### 2017-06-04 paper/figures/example-stats.tex
words about transit and more steps for transit

\begin{floattable}
\begin{deluxetable}{cccccccc}
\caption{The computational cost and convergence stats for each example. \label{tab:example-stats}}
\tablehead{& \colhead{$N$} & \colhead{$J$} & \colhead{direct\tablenotemark{a}} & \colhead{\celerite\tablenotemark{b}} & \colhead{dimension\tablenotemark{c}} & \colhead{evaluations\tablenotemark{d}} & \colhead{$N_\mathrm{eff}$\tablenotemark{e}}\\
&&& \colhead{ms} & \colhead{ms} &&&}
\startdata
1 & 200 & 1 & 2.85 & 0.26 & 3 & 80000 & 1737 \\
2 & 100 & 2 & 0.67 & 0.36 & 4 & 80000 & 1134 \\
3 & 6950 & 2 & 8119.11 & 1.47 & 4 & 176000 & 2900 \\
4 & 1440 & 8 & 828.93 & 2.74 & 9 & 640000 & 1443 \\
5 & 1000 & 2 & 88.30 & 0.96 & 13 & 1280000 & 1490 \\
\enddata

\tablenotetext{a}{The computational cost of computing the GP model using the
general Cholesky factorization routine implemented in the Intel MKL.}
\tablenotetext{b}{The computational cost of computing the GP model using
\celerite.}
\tablenotetext{c}{The total number of parameters in the model.}
\tablenotetext{d}{The total number of evaluations of the model to run MCMC to
convergence.}
\tablenotetext{e}{The effective number of independent samples estimated by
computing the integrated autocorrelation time of the chain.}
\end{deluxetable}
\end{floattable}