### 2017-06-12 paper/response.txt
adding response

We would first like to thank the referee for their thorough and constructive
review. We have taken the feedback to heart and we think that the manuscript
has been greatly improved as a result.

Beyond addressing each of the referee's concerns in detail, we have also
derived a new algorithm for solving celerite models that is about a factor of
20 faster than the submitted version (and about and order of magnitude faster
than a reference implementation of a Kalman filter based method). We have
updated the manuscript to describe and test this new method in Section 5. We
hope that the referee will consider this updated method and welcome any
feedback.

The only other major structural change was to switch the order of the section
titled "celerite as a model of stellar variations" and the section describing
the algorithm. "Stellar variations" is now Section 4 and the implementation is
discussed in Section 5.

Below we address each of the referee's comments in detail:

## Major Issues

> Unclear model specification: A Bayesian model is composed of the likelihood
> function and joint prior distribution of the unknown parameters. The likelihood
> function of the unknown parameters, L(θ,α), for the proposed GP method is
> implicitly set to Equation (3). First of all, the components of the two
> unknown vectors, θ and α, are not the same for all of the examples, and thus
> the authors must define these clearly in each example. For example, two vectors
> are θ = (?, ?, ?) and α = (?, ?, ?) so that the likelihood function is uniquely
> determined. Next, the authors must clearly specify the joint prior distribution
> p(θ,α). If the authors use a Uniform prior distribution, please clearly specify
> the range of the Uniform distribution because there are uncountably many
> Uniform prior distributions depending on the range. Once the likelihood
> function and joint prior distribution are clearly specified, then the joint
> posterior distribution, p(θ, α | X, y), is uniquely determined and I cannot say
> that ‘your model specification is unclear.’ Please clearify your Bayesian
> models in all the numerical illustrations including the two simulations and
> three realistic data analyses. It may be possible that the full posterior
> distribution may contain unknown parameters more than θ and α in their
> numerical illustrations, e.g., nuisance parameters in Section 7.1. Again,
> please follow the principle rule by specifying the likelihood function of all
> the unknown parameters and their prior distributions. Please do not use any
> MCMC method before clearly specifying the full posterior distribution.
> Otherwise users do not know what the distribution of the posterior sample
> represents.

We have updated each example to include a list of each parameter and the chosen
priors. We also include a table for each example that lists this information.

> Posterior propriety: Posterior propriety means that your full posterior
> distribution has a finite integral,
> [...]
> This is the basis of all the Bayesian analyses because the fundamental Bayes
> rule holds if and only if posterior propriety holds. Treating X and y as the
> observed data, we can derive the Bayes
> rule as follows.
> [...]
> Without posterior propriety, the joint posterior distribution p(θ,α | X,y) is
> not guaranteed to be a probability distribution because the denominator in (2)
> blows up. Also, without posterior propriety, there is no way to know whether a
> Markov chain converges to the correct target (stationary) posterior
> distribution, p(θ, α | X, y). This is because a Markov chain is null recurrent
> for an improper posterior distribution and thus the stationary distribution of
> a Markov chain does not exist (see the article in footnote 1 for details). If
> the authors use any improper prior distributions, e.g., for a parameter θi,
> log(θi) ∼ Uniform(−∞, ∞) or θi ∼ Uniform(0, ∞), then the authors must mention
> whether posterior propriety holds before any MCMC method is used. (The authors
> may say that their MCMC method produces reasonable posterior samples of the
> unknown parameters, recovering their true values correctly. This is not about
> whether the MCMC method recovers the true values or not. This is about whether
> the Markov chain converges to the targeted stationary distribution. Your target
> posterior distribution may or may not contain the true values and recovering
> true values does not necessarily mean that your chain converges to the targeted
> posterior distribution. Converging to the stationary distribution and
> recovering true values are two different problems.) A well-known insidious
> property of posterior propriety is that the MCMC sampler may appear to work
> well even when the posterior distribution is improper, and thus practitioners
> do their sciences based on non-existent posterior distributions. There are only
> two ways to guarantee posterior propriety: (i) Using a jointly proper prior
> distribution, p(θ,α) dθdα < ∞, guarantees posterior propriety, which will
> become clear when the authors clarify their Bayesian models. (ii) Showing that
> (1) holds in a mathematically rigorous way.

In each example, we were already using (unspecified!) proper priors. We have
updated each example to include a discussion of the priors and these are fully
specified in the tables included with each example.

> Model checking: The authors provide a method for a model fitting but I do not
> know why I trust their estimates or posterior distributions. In each example,
> there should at least be graphical model checking, e.g., by plotting residuals
> with respect to an optimal model. This is one way to conduct posterior
> predictive check, i.e., generating simulated data using the proposed model
> given the posterior samples of the parameters and visually checking whether
> these simulated data match well with the observed data or whether the residuals
> do not have any patterns. Successful model checking could support an argument
> that, at least, the present model is not significantly misspecified, so that
> the uncertainty estimates can be trusted.

We have updated the figures in each example to include plots of the residuals
between the data and the maximum a posteriori model.

> Unknown data size: In Sections 6 and 7, the authors never specify the data size
> in each example. It is quite shocking to me that Section 7.1 even does not
> display any data at all. Section 6 does not seem to motivate the argued
> ‘scalability’ because the data size appears to be small in Figures 6 and 7. Are
> the authors motivated by a realistic problem to develop a scalable GP?  Why do
> the authors need a scalable GP? Do we really need the argued ‘scalability’ to
> fit the data used in each example? Are there any examples that motivate this
> research on the scalability of the GP with the specific covariance function?
> Throughout the manuscript, the authors argue that a conventional GP is not
> suitable for large data sets. Then please show me or readers a realistic
> example (evidence) where a conventional GP is almost impossible to implement
> while celerite works well. I feel that a conventional GP with the same
> covariance functions (but based on a matrix inversion, i.e., without the trick
> for the scalability) may work well at least for the simulation studies; I
> cannot judge whether my feeling is correct for the realistic examples because
> the authors do not clarify the data sizes. If my feeling is wrong, then please
> show me evidence.

We agree that this discussion was severely lacking from the submitted
manuscript and we have included a discussion of the need for scalable methods
in the introduction, in each example, and in a summary subsection at the end of
Section 6. We now fully specify the data sizes for each example and discuss the
practicality of applying general direct solvers in each case.

> Tightening descriptions is needed.
> - Abstract: “Gaussian Processes are a popular class of models used for this
>   purpose but, since the computational cost scales as the cube of the number of
>   data points, their application has been limited to relatively small
>   datasets.” Is it true for all GPs? This sentence gives me an impression that
>   all GPs scale as the cube of the number of data points. If I use a
>   Kalman-filtering-based GP, then what can the authors say? Please tighten the
>   sentence. The authors are posing a problem of GPs that involve a (large)
>   matrix inversion.
>
> - Abstract “but we also demonstrate that it is effective in many other cases.”,
>   Section 2 “improving this scaling in many circumstances” and “many
>   astronomical data analysis problems”: The authors provide three realistic
>   examples and in Section 9 (Summary) the authors say that the method is
>   expected to work well in many other situations. Expectations are
>   expectations. I am not sure whether the authors can use the word ‘many’ such
>   often based on three realistic examples. Do these three examples represent
>   various astronomical data analyses listed in Section 9?
>
> - Abstract: “flexible, fast, and most importantly, interpretable” The
>   covariance function for celerite is restrictive as the authors admit in
>   Section 8. I am not sure whether the word ‘flexible’ describes celerite in an
>   objective way.
>
> - Section 1: “However, there is no further constraint on the data or the
>   model.” This sentence is wrong because the authors mention that the
>   covariance function must be stationary on page 4.
>
> - Section 1: “As we will discuss in the following pages, this choice of kernel
>   function is not as restrictive as it first seems because it arises naturally
>   in some physical systems and it can be used as an effective model in many
>   other cases.” This sentence is vacuous due to the words, ‘seems’, ‘some’, and
>   ‘many’. As for the word ‘seems’, the covariance function that celerite allows
>   is restrictive at my first, second, and third glances and the authors also
>   admit it in Section 8. As for the word ‘some’, the three examples handles
>   three physical systems. As for the word ‘many’, in Section 9 (Summary) the
>   authors say that the method is expected to work well in many other
>   situations. But I keep feeling that this kind of argument, ‘in many other
>   cases’, based only on their expectations is appropriate.
>
> - Section 1: “GP models have been widely used across the physical sciences but
>   their application is generally limited to small datasets” How small is small
>   to the authors? A light curve with 999 data points? Please clarify the cases
>   where the authors think the data size is small so that the readers can judge
>   how small is small. Do the authors use a ‘large’ data set in the examples to
>   show that a typical GP with a matrix inversion never works while celerite
>   successfully works?

For each of the above, we agree and have address each concern.

> - Section 6: “Even in these cases, GPs can be useful
>   effective models and celerite provides computational advantages over other GP
>   methods.” This may imply that celerite has a computational advantage over any
>   GP methods. The authors mention in Section 8 that a Kalman-filtering is in
>   practice somewhat faster and the memory requirements are smaller. So the
>   sentence is not correct.

The current method is now faster than Kalman-filtering methods so this
statement is now true.

> - Section 7.1: “initialize our celerite model using
>   a grid search in the parameter space” and “This model requires about a minute
>   of computation time to find the maximum likelihood parameters and then it
>   takes about an hour to run the MCMC sampling to convergence on a dual-CPU
>   laptop.”: For the first sentence, I have never read in the previous sections
>   about how to initialize the model using a grid search in the parameter space.
>   What do the authors actually do? Why do the authors do the grid search? What
>   is the dimension of the parameter space and the grid? The second sentence is
>   also surprising because the method description in the previous sections never
>   says that the implementation procedure is actually a two-step procedure. I
>   guess that the authors use MLEs as initial values of their MCMC method.
>   Clearly specify the entire procedure including how to set up initial values
>   of the celerite model parameters for an MCMC in the section that describes
>   the method implementation. In the numerical illustrations, the authors are
>   not supposed to say anything new or unexplained about the implementation.

We have updated the discussion of each example to include a description of the
initialization procedure.

> - Section 7.1: “another GP solver” This phrase is broad. Which GP solver do the
>   authors indicate? Please be specific.

Fixed.

## Minor concerns

> Abstract: “a comparison to existing scalable Gaussian Process methods.” The
> authors clearly mention that any CARMA(p, q) model can be matched to their
> celerite model in Section 4.2. Then why not numerically compare the scaleable
> method of fitting CARMA(p, q) with celerite?

We have added an explicit numerical comparison between Kalman-filtering based
CARMA models and celerite. Figure 13 has been added to show the computational
scaling of the Kelly et al. (2014) solver and find that it is about a factor of
10 slower than celerite.

> Section 2: The authors say that the log likelihood function is Equation (3).
> However, Equation (3) is clearly a function of y given X, θ, and α.

We have fixed the sloppy notation by defining this function as L(θ,α).

> Until Section 4, I have thought that the authors may be using a
> Kalman-filtering for a specific form of a covariance matrix. Maybe it is my
> fault. When a scalable GP with O(n) is mentioned, the first thing I think about
> is a Kalman-filtering. What about mentioning that the proposed method is not a
> Kalman-filtering but is also scalable with O(n) in the early stage of the
> manuscript.

Good point. We have added this clarification to the introduction.

> Section 2: ‘using a non-linear optimization routine’ I am just curious about
> how Nocedal and Wright (2006) obtain the uncertainty of their point estimate
> because I believe that point estimation can be misleading without uncertainty
> quantification.

We absolutely agree! We have emphasized the fact that this estimate would only
be a point estimate.

> Section 2: Markov Chain Monte Carlo → Markov chain Monte Carlo. The word
> ‘chain’ is not a name of a person.

Fixed.

> The parameter J is an obvious tuning parameter of the proposed method. I was
> surprised when I read ‘but, in many cases, small values of J are sufficient” in
> Section 9. Maybe the authors think it is ok in an astronomical journal to give
> users a broad and unclear guideline on how to set the tuning parameter. This
> issue would be a major issue if it were for a publication in a statistical
> journal. Please keep in mind that without setting J no one can implement
> celerite. Imagine a user who wants to use celerite but do not know how to set
> J. Are the authors going to say ‘use a small value of J’? The current sentence
> sounds very irresponsible. This sentence gives me an impression that J = 1 must
> work well ‘in many cases,’ although I am not sure how many is many. The most
> important thing is to specify what made the authors believe that the small
> values of J would work well in many cases so that users can judge whether the
> small values of J are also appropriate for their sciences. Please show the
> readers the evidence for the argument or a clear procedure to determine J so
> that they can follow it for their data analyses. If the authors do not have any
> evidence or a procedure, i.e., the choices of J in the examples of the current
> manuscript are all ad-hoc, please do more research on it. Please do not use the
> word ‘small’ because this word means nothing to users. Also please specify the
> tuning procedure in the section regarding the implementation because without
> setting J no one can implement celerite. Finally, please clarify how the
> authors choose J in each of two simulation studies and three realistic data
> analyses, which will be helpful for users.

We respectfully disagree with the referee's interpretation of J as a "tuning
parameter" of the method and instead argue that this is just the general
problem of model selection when using GPs. Every GP analysis must specify a
kernel and celerite is no different. That being said, we agree with the
concerns about the vague statement about "small J" so we have removed this and
added a section (5.4) discussing options for model selection with some
celerite-specific recommendations.  The models in all examples except Example
2 are chosen using physical motivation and we have updated Example 2 to
include a discussion and a new figure to explain the procedure that we used to
select the model in that case.

> Section 3: The mixture idea shown in Equation (6) also appears in the
> literature. The discussion about this idea and the one in Ambikasaran (2015)
> may be interesting but can be distracting. Please give enough credit for the
> earlier work, if any.

Added.

> Section 3: Ornstein-Uhlenbeck process may need a citation.

Done.

> After Equation (19) in Section 4.1: There are not b and d in Equation (7).
> Do the authors mean bj and dj?

This section has been removed and replaced by our new derivation.

> Section 5: “This suggests that the Matern-3/2 covariance can be approximated
> using the celerite framework with a small value of f in Equation (54) but we
> caution that this might lead to numerical issues for the solver.” In the
> previous description, there is no approximation at all. Where does the
> approximation occur in the previous description? This sentence also sounds
> problematic. It indicates that some models definitely within the celerite
> framework may not be feasible because of numerical issues. Then the authors
> must specify feasible class of models among all the possible models within the
> celerite framework in order not to disappoint users.

We have expanded our discussion in this section and made the approximation
explicit. The numerical issue comes from the fact that the parameter b_j is
infinite for Q = 1/2. We now state this explicitly in the manuscript.

> Section 6.2: This is an impressive result. However, at the same time, I feel
> that this is a dangerous result because the proposed model does not reveal any
> evidence on a potential model mis-specification.

As with all the examples, we have added a graphical model check. As we discuss
in the text, any broad use of this method would require more research to
determine the detailed limits of applicability, but we think that that is
beyond the scope of this paper.

> Section 7.1: “is is” is a typo.

Fixed.

> Section 8: Comparison to a 20 year old method that the current method is
> exactly built upon? Without reading further, I can imagine that the proposed
> method is better.

We have removed this comparison.

> Section 8: Comparison to CARMA. The authors say that any CARMA model has a
> corresponding celerite model. This means that I can fit any CARMA model via
> Kalman-filtering, obtain posterior samples of the unknown parameters of the
> CARMA model, match these posterior samples of the CARMA parame- ters with
> corresponding celerite parameters, and take advantage of the better physical
> interpretability of the celerite parameters. Why should the readers use
> celerite at the beginning instead of using Kalman-filtering in fitting a CARMA
> model? I am not quite motivated at this point.

We have expanded our discussion in this section to provide further motivation.

> Section 9: The authors list all the possible cases where celerite may work
> well. I think this section (summary) is full of expectations (too many
> expectations for a scientific journal in my opinion). I want to understand why
> the authors expect celerite to work well in such many cases. If the authors
> have clear evidence for these cases, then what about making another section for
> potential applications? In my opinion if this section is for ‘Summary’, then it
> is not a good place to talk about things that have not explained in the
> previous sections.

We agree and we have changed the section title to "Discussion". We have removed
some speculation from this section and added a subsection where we explicitly
discuss some potential future applications.